---
title: "BTB"
output:
  md_document:
    variant: markdown_github
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F)
```

```{r load_packages}
library("lubridate")
library("tidyverse")
library("knitr")
```

```{r}
nF<-function(x) {ifelse(is.na(x), F,x)}
countN<-function(x) {as.integer(sum(!is.na(x)))}
rmean<-function(x){mean(x,na.rm=T)}
rmeanr<-function(x){round(mean(x,na.rm=T),2)}
rmedian<-function(x){median(x,na.rm=T)}
rsum<-function(x) {sum(x,na.rm=T)}

```


```{r load_data}
load(file = "/Users/srhrnkn/Documents/authors/btb.Rdata")
```


```{r create_objects}
#summarize by interviewee: counts by gender, percents, bday, age
byint<-btb %>% group_by(interviewee.GRname,interviewee.id, interviewee.gender.use,interviewee.birthdate,Date) %>% summarise(countF=rsum(author.gender.use=="f"),countM=rsum(author.gender.use=="m"),countauth=countN(author.gender.use=="f")) %>% mutate(percF=countF/countauth,age=as.period(interval(interviewee.birthdate,Sys.Date()))$year) %>% mutate(varf=(1-percF)/countauth) %>% arrange(percF)


#summarize by author: mentiongs, bday, age
byauth<-btb %>% group_by(author.GRname, author.gender.use,author.birthdate) %>% summarise(countFint=rsum(interviewee.gender.use=="f"),countMint=rsum(interviewee.gender.use=="m"),countment=countN(interviewee.gender.use)) %>% mutate(percF=countFint/countment,age=as.period(interval(author.birthdate,Sys.Date()))$year) %>% arrange(desc(countment))

#authors with more than three mentions by date of interview
highfreqauthbydate<-btb %>%  filter(author.GRname %in% byauth$author.GRname[byauth$countment>3]) %>% select(author.GRname, interviewee.GRname, Date, interviewee.gender.use) %>% left_join(byauth[,c("author.GRname","countment")]) %>% arrange(desc(countment),author.GRname,Date) %>% mutate(author=factor(author.GRname,levels = rev(unique(author.GRname))))

#colors for plots
plotcolors<-data.frame(abbr=c("f","m"),adj=c("female","male"),noun=c("women","men"),color=c("#a04646","#4661a0"),stringsAsFactors = F)
oranges<-colorRampPalette(c("grey","orange"))
sixoranges<-oranges(6)
```


## New York Times "By The Book" Interviews  
Interview dates: **`r paste0(as.character(month(min(byint$Date),label =  T,abbr = F))," ",day(min(byint$Date)),", ",year(min(byint$Date))," - ",as.character(month(max(byint$Date),label =  T,abbr = F))," ",day(max(byint$Date)),", ",year(max(byint$Date)))`**  

Total interviews: **`r nrow(byint)`**  

Total unique authors mentioned: **`r nrow(byauth)`**  
  
The New York Times Book Review's "By the Book" feature is a weekly interview exploring reading habits past, present, and planned. The interviews yield author recommendations - an average of **`r round(nrow(btb)/nrow(byint),)`** authors mentioned by name in each interview. The dataset analyzed below is a result of code that looks at the text of each interview, identifies strings of text that look like author names, and verifies them by looking for corresponding names in the Goodreads author database. The gender and birthdate information comes from the data in the Goodreads database, with missing data filled in from the authors' Wikipedia pages.

### Gender

Male interviewees are more likely to mention male authors.
  
  
  
  

```{r }
#new
btb %>% group_by(interviewee.gender.use) %>% count(author.gender.use) %>% mutate(grcount=sum(n),perc=n/grcount,percformat=paste0(100*round(perc,2),'%')) %>% ggplot(aes(x=interviewee.gender.use,y=n,fill=author.gender.use)) + 
  geom_bar(stat = "identity",position = "fill",width = .5) + geom_text(aes(label=paste0(percformat,"\n(",n,")")), position=position_fill(vjust=.5))+
  scale_fill_manual(values = plotcolors$color,name="Authors mentioned: ",labels=plotcolors$noun,guide=guide_legend(reverse = T))  + scale_x_discrete(labels=paste0(plotcolors$adj," interviewees (",c(sum(byint$interviewee.gender.use=="f"), sum(byint$interviewee.gender.use=="m")),")")) + 
  theme(axis.title=element_blank(),axis.text.x = element_blank(),axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),panel.background = element_rect(fill = NA),legend.direction = "horizontal",legend.position = "top",legend.justification = c(0, 0)) + 
  coord_flip() #+
  #theme(aspect.ratio = .3,plot.margin = unit(c(0,0,0,0),units = "pt"))



```
  
  
  
  
But there is plenty of variation among interviewees in the total number of authors mentioned & in the gender split. 
  
  
  
  
```{r}
ggplot(byint %>% mutate(intervieweelevels=factor(interviewee.GRname,levels=byint$interviewee.GRname[order(byint$countauth)])),  aes(x = intervieweelevels)) +
         geom_bar(aes(y=byint$countF),stat = "identity",fill=plotcolors$color[1]) +
         geom_bar(aes(y=-1*byint$countM),stat = "identity",fill=plotcolors$color[2]) +
  xlab("Interviewees") + ylab("Author mentions") +
  scale_x_discrete(labels=paste(byint$interviewee.GRname[order(byint$countauth)],byint$countauth[order(byint$countauth)])) +
  scale_y_continuous(breaks = seq(-40, 40, 20),
                     labels = as.character(abs(seq(-40, 40, 20)))) +
  coord_flip()
```
  
  
  
  
Older interviewees are both more likely to be male (**`r 100*rmeanr(byint %>% filter(interviewee.gender.use=="m") %>%  select("interviewee.birthdate") %>% transmute(old=year(interviewee.birthdate)<1960) %>% pull())`%** of male interviewees were born before 1960, vs **`r 100*rmeanr(byint %>% filter(interviewee.gender.use=="f") %>%  select("interviewee.birthdate") %>% transmute(old=year(interviewee.birthdate)<1960) %>% pull())`%** of female interviewees) and more likely to have recommended a lower percentage of female authors. 
  
  
  
  
```{r warning=F}
ggplot(byint,aes(x=interviewee.birthdate,y = percF,color=interviewee.gender.use,size=countauth)) +geom_point() +scale_color_manual(values=plotcolors$color, name="interviewee\ngender") + scale_size(name="authors\nmentioned") + ylab("Female authors mentioned (percent)") + xlab("Interviewee birthdate") +scale_y_continuous(labels = function(x){as.character(100*x)},limits = c(0,1))
```




###Frequently mentioned authors

This data has a long tail - there are **`r nrow(byauth)`** unique authors mentioned across the **`r nrow(byint)`** interviews, and **`r byauth %>% filter(countment==1) %>% nrow()`** of them are only mentioned once. The most frequently mentioned authors are mentioned by a fifth of all interviewees. The chart below shows the most frequently mentioned authors, the number of times they were mentioned, and the dates they were mentioned. 
  
  
  
  
```{r}
#higher frequency by date mentioned
ggplot(highfreqauthbydate,aes(x=Date,y=author)) + 
  geom_point(aes(color=interviewee.gender.use)) + 
  theme(legend.position="top",legend.direction = "horizontal",legend.justification = c(0, 0),axis.title=element_blank(),axis.ticks = element_blank(),plot.title = element_text()) + 
  scale_y_discrete(labels=rev(unique(paste0(highfreqauthbydate$author," (",highfreqauthbydate$countment,")")))) +  
  scale_x_date(date_minor_breaks = "1 month") +
  scale_color_manual(values=plotcolors$color,name="Date mentioned; gender of interviewee:") 
```


```{r}
#make df
heatmapint <- btb %>%  filter(author.GRname %in% byauth$author.GRname[byauth$countment>0]) %>% select(author.GRname, interviewee.GRname, Date, interviewee.gender.use) %>% left_join(byauth[,c("author.GRname","countment")]) %>% arrange(desc(countment),author.GRname,Date) %>% mutate(author=factor(author.GRname,levels = rev(unique(author.GRname))))
heatmapint <- full_join(x = heatmapint[,c("author","interviewee.GRname")],y = heatmapint[,c("author","interviewee.GRname")],by = "author" )
#rearrange columns and delete entries where x and y are same 
heatmapint$Weight<-1
#heatmapint$Weight[which(heatmapint$interviewee.GRname.y==heatmapint$interviewee.GRname.x)]<-NA

#set upper half to NA (getting rid of this, it's nice to be able to see full rows for everyone)
#heatmapint$Weight[match(heatmapint$interviewee.x, heatmaplevels)<match(heatmapint$interviewee.y, heatmaplevels)]<-NA

#data for plot - sum weight by interview mentions
heatmapintplot<-heatmapint %>% group_by(interviewee.GRname.x,interviewee.GRname.y) %>% summarize(Weight=sum(Weight))
#create levels to sort in order of most appearances (pull turns this back into a character vector)
heatmapintlevels<-heatmapintplot  %>% group_by(interviewee.GRname.x) %>% summarize(n=max(Weight)) %>% arrange(desc(n)) %>% select(interviewee.GRname.x) %>% transmute(levels=as.character(interviewee.GRname.x)) %>% pull()
heatmapintplot$interviewee.x<-factor(heatmapintplot$interviewee.GRname.x,levels = heatmapintlevels)
heatmapintplot$interviewee.y<-factor(heatmapintplot$interviewee.GRname.y,levels = heatmapintlevels)
#pull out total authors mentioned as separate character var to plot on diagonal
heatmapintplot$authtotal<-as.character(heatmapintplot$Weight)
heatmapintplot$authtotal[which(heatmapintplot$interviewee.y!=heatmapintplot$interviewee.x)]<-NA
heatmapintplot$Weight[which(heatmapintplot$interviewee.y==heatmapintplot$interviewee.x)]<-NA
```
  
  
  
  
Below we can see instances where two interviewees mentioned the same author. Note that although some interviewees mentioned dozens of authors, the maximum overlap in authors mentioned between any two interviewees was **`r max(heatmapintplot$Weight,na.rm=T)`**.
  
  
  
  
```{r fig.width = 9, fig.height = 9 }
ggplot(heatmapintplot, aes(x = interviewee.x,y = interviewee.y)) + 
  geom_tile(aes(fill = as.factor(Weight)),color="grey") +  
  theme(axis.line = element_blank(), 
        panel.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_fill_manual(values = c(sixoranges,"white"),guide = guide_legend(title = NULL)) +
  coord_fixed() +
  #geom_text(data = heatmapintplot$Weight[which(heatmapintplot$interviewee.x==heatmapintplot$interviewee.y)] ) +
  geom_text(aes(label=authtotal),na.rm = T) +
  ggtitle(label = "Count of authors mentioned in common by each pair of interviewees\n(Diagonal shows total authors mentioned by that interviewee)")
```
  
  
  
  
We can also see how many times pairs of authors were mentioned together in the same interview.
  
  
  
  
```{r fig.width = 8, fig.height = 8 }

#make df
heatmapauth <- btb %>%  filter(author.GRname %in% byauth$author.GRname[byauth$countment>3]) %>% select(author.GRname, interviewee.GRname, Date, interviewee.gender.use) %>% left_join(byauth[,c("author.GRname","countment")]) %>% arrange(desc(countment),author.GRname,Date) %>% mutate(author=factor(author.GRname,levels = rev(unique(author.GRname))))
heatmapauth <- full_join(x = heatmapauth[,c("author","interviewee.GRname")],y = heatmapauth[,c("author","interviewee.GRname")],by = "interviewee.GRname" )
#rearrange columns and delete entries where x and y are same 
heatmapauth$Weight<-1
#heatmapauth$Weight[which(heatmapauth$author.y==heatmapauth$author.x)]<-NA
heatmaplevels<-heatmapauth  %>% group_by(author.x) %>% tally() %>% arrange(desc(n)) %>% select(author.x) %>% transmute(levels=as.character(author.x)) %>% pull()
heatmapauth$author.x<-factor(heatmapauth$author.x,levels = heatmaplevels)
heatmapauth$author.y<-factor(heatmapauth$author.y,levels = heatmaplevels)
#create levels to sort in order of most appearances (pull turns this back into a character vector)


#set upper half to NA (getting rid of this, it's nice to be able to see full rows for everyone)
#heatmapauth$Weight[match(heatmapauth$author.x, heatmaplevels)<match(heatmapauth$author.y, heatmaplevels)]<-NA

heatmapauthplot<-heatmapauth %>% group_by(author.x,author.y) %>% summarise(Weight=sum(Weight))

heatmapauthplot$mentiontotal<-as.character(heatmapauthplot$Weight)
heatmapauthplot$mentiontotal[which(heatmapauthplot$author.y!=heatmapauthplot$author.x)]<-NA
heatmapauthplot$Weight[which(heatmapauthplot$author.y==heatmapauthplot$author.x)]<-NA


ggplot(heatmapauthplot, aes(x = author.x,y = author.y)) + 
  geom_tile(aes(fill = as.factor(Weight)),color="white") + 
  theme(axis.line = element_blank(), 
        panel.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_fill_manual(values = c(sixoranges[1:5],"white"),guide = guide_legend(title = NULL)) +
  coord_fixed() +
  ggtitle(label = "Top authors: count of interview overlap\n(Diagonal shows total times author was mentioned)") +
  geom_text(aes(label=mentiontotal),na.rm = T)

```


