---
title: "BTB"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F)
```

```{r load_packages}
library("lubridate", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("tidyverse", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("knitr", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
```

```{r}
nF<-function(x) {ifelse(is.na(x), F,x)}
countN<-function(x) {as.integer(sum(!is.na(x)))}
rmean<-function(x){mean(x,na.rm=T)}
rmeanr<-function(x){round(mean(x,na.rm=T),2)}
rmedian<-function(x){median(x,na.rm=T)}
rsum<-function(x) {sum(x,na.rm=T)}
```


```{r load_data}
load(file = "/Users/srhrnkn/Documents/authors/btb.Rdata")
```


```{r create_objects}
#summarize by interviewee: counts by gender, percents, bday, age
byint<-btb %>% group_by(interviewee.GRname,interviewee.id, interviewee.gender.use,interviewee.birthdate,Date) %>% summarise(countF=rsum(author.gender.use=="f"),countM=rsum(author.gender.use=="m"),countauth=countN(author.gender.use=="f")) %>% mutate(percF=countF/countauth,age=as.period(interval(interviewee.birthdate,Sys.Date()))$year) %>% mutate(varf=(1-percF)/countauth) %>% arrange(percF)


#summarize by author: mentiongs, bday, age
byauth<-btb %>% group_by(author.GRname, author.gender.use,author.birthdate) %>% summarise(countFint=rsum(interviewee.gender.use=="f"),countMint=rsum(interviewee.gender.use=="m"),countment=countN(interviewee.gender.use)) %>% mutate(percF=countFint/countment,age=as.period(interval(author.birthdate,Sys.Date()))$year) %>% arrange(desc(countment))

#authors with more than three mentions by date of interview
highfreqauthbydate<-btb %>%  filter(author.GRname %in% byauth$author.GRname[byauth$countment>3]) %>% select(author.GRname, interviewee.GRname, Date, interviewee.gender.use) %>% left_join(byauth[,c("author.GRname","countment")]) %>% arrange(desc(countment),author.GRname,Date) %>% mutate(author=factor(author.GRname,levels = rev(unique(author.GRname))))
```


New York Times By The Book interviews, `r paste0(as.character(month(min(byint$Date),label =  T,abbr = F))," ",day(min(byint$Date)),", ",year(min(byint$Date))," - ",as.character(month(max(byint$Date),label =  T,abbr = F))," ",day(max(byint$Date)),", ",year(max(byint$Date)))`

Total interviews: `r nrow(byint)`  
Total unique authors mentioned: `r nrow(byauth)`  

Top 21 authors mentioned:  
```{r}
byauth %>% data.frame() %>% select(author=author.GRname,mentions=countment) %>% slice(1:21) %>% kable(align="lc",format = "html",table.attr = "style='width:30%;'")
```

Authors mentioned by gender of author & interviewee: 

```{r fig.width=3}
#authormentions by gender raw counts
ggplot(btb,aes(x=interviewee.gender.use,fill=author.gender.use)) + geom_bar() + xlab("Interviewee gender") + ylab("authors mentioned")  + scale_fill_manual(values = c("darkgrey","lightgrey"),name="Author gender",labels=c("female","male")) + theme_bw() + scale_x_discrete(labels=paste0(levels(btb$interviewee.gender.use)," (",c(sum(byint$interviewee.gender.use=="f"), sum(byint$interviewee.gender.use=="m")),")"))


#percentages

ggplot(data.frame(sweep(x = table(btb$author.gender.use,btb$interviewee.gender.use,dnn = c("authors","interviewees")),MARGIN = 2,STATS = margin.table(table(btb$author.gender.use,btb$interviewee.gender.use),2),FUN = "/")),aes(x=interviewees,y=Freq,fill=authors)) + geom_bar(stat = "identity") + xlab("Interviewee gender") + scale_fill_manual(values = c("darkgrey","lightgrey"),name="Author gender",labels=c("male","female")) + theme_bw() + scale_x_discrete(labels=paste0(levels(btb$interviewee.gender.use)," (",c(sum(byint$interviewee.gender.use=="f"), sum(byint$interviewee.gender.use=="m")),")"))
```

Counts of author mentions by interviewee

```{r}
ggplot(btb %>% mutate(intervieweelevels=factor(interviewee.GRname,levels=byint$interviewee.GRname[order(byint$countauth)])), aes(intervieweelevels)) + geom_bar(aes(fill = author.gender.use))  + theme(legend.position="none") + xlab("interviewee") + coord_flip()
```

Most mentioned authors by date mentioned

```{r}
#higher frequency by date mentioned
ggplot(highfreqauthbydate,aes(x=Date,y=author)) + geom_point(aes(color=interviewee.gender.use)) + theme(legend.position="none") #+ scale_y_discrete(breaks=1:length(levels(highfreqauthbydate$author)),labels=unique(paste0(highfreqauthbydate$author," (",highfreqauthbydate$countment,")")))
```

